Microxanox - an R package for simulating an $MICR$obial ecosystem that can occupy $OX$ic or $ANOX$ic states.

### 2022-05-02 inst/manuscript/manuscript_microxanox.Rmd
New Structure next step

The aims of the `microxanox` package are twofold. Firstly, the package aims at reproducing the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED). Secondly, to take these results one step further, it includes new functionality to address our research question as presented in [@REF_NEEDED].

@@

The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED]. Consequently, any more substantial changes and adaptations, are likely to need a change in the source code.

Nevertheless, the model is structured in a way which builds on a modular structure, so that e.g. the event definition can easily be changed, or other aspects can be adjusted. All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.

@@

1. all parameter needed to run a simulation or find a stable state are contained in a single parameter object. This object is created by using the functions `new_..._parameter()`, `new_initial_state()` and `new_strain_parameter()`. Which one of the `new_..._parameter()` functions has to be used when, will be discussed in the section  \@ref(runsim)  and in the [User Guide](LINK_NEEDED).
2. The function call `run_...(parameter)` will run the simulation using the parameter as defined in the object `parameter`.
3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additionl slot named `results` which contains the results of the run
4. As this return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.

@@

## Software Functionalities
> {>>Present the major functionalities of the software.<<}

We will discuss the general structure and functionality of the package without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).

@@

# Illustrative Examples
> {>>Present the major functionalities of the software.<<}

@@

The individual simulation (`run_simulation()` function) is the working horse in this package. In this function, the ODEs are solved. The function needs only one argument - an object as created by the function `new_runsim_parameter()`. One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`. For a detailed description of the parameter and how they are created please see the User Guide which acomnpanies the package or is available at [User Guide](@LINK_NEEDED)

@@

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This functio is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in \@ref(fig:plot-dynamics).

@@

This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()` and a number indicating the .

@@

The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes. The third is taken from the @REF_NEEDED for which this R package was designed. All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.

## Regime shifts during temporal environmental change

@@

## The extent of hysteresis depends on community composition

One of the reasons to develop this package was to reproduce the results presented in @Bush2017, this was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED). All aspects in the paper could be reproduced and are shown in the vignette.

@@

## Effects of functional diversity on regime shifts

@@

>>>>

The aims of the `microxanox` package are twofold. Firstly, the package aims at reproducing the results shown by [@Bush2017], which is accomplished in the [vignette Partial reproduction of Bush et al](LINK NEEDED). Secondly, to take these results one step further, it includes new functionality to address our research question as published in [@REF_NEEDED].

@@

We will discuss the general structure and functionality of the package here without going into to much detail. A more detailed discussion can be found in the [User Guide](LINK_NEEDED).

The package is not intended to provide a modelling framework which can be adjusted easily to all needs, but primarily a tool to implement the model used by [@Bush2017] and to extend it to our needs [@REF_NEEDED]. Consequently, any more substantial changes and adaptations are likely to need a change in the source code.

Nevertheless, the source code is structured in modular, so that e.g. the event definition can easily be changed, or other aspects can be adjusted. All values in the parameter object can be changed as needed and the general structure of the code should make it not to difficult to adapt the model to other similar systems.

@@

1. all parameter needed to run a simulation or find a stable state are contained in a single parameter object. This object is created by using the functions `new_..._parameter()`, `new_initial_state()` and `new_strain_parameter()`. Which one of the `new_..._parameter()` functions has to be used when, will be discussed in the section  \@ref(runsim) and in more detail in the [User Guide](LINK_NEEDED).
2. The function call `run_...(parameter)` will run the simulation using the parameter as defined in the object `parameter`.
3. The return value of the `run_...(parameter)` function is identical to the parameter object plus an additional slot named `results` which contains the results of the run
4. As the return value contains all parameter, it is possible to re-run the simulatuion by simply running `run_...(result)`.

@@

@@

@@

The individual simulation (`run_simulation()` function) is the working horse in this package. In this function, the ODEs are solved. The function needs only one argument - an object as created by the function `new_runsim_parameter()`. One parameter of this object is the `strain_parameter` which can be created by the function `new_strain_parameter()`. For a detailed description of the parameter, their meaning and how they are created please see the User Guide which accompanies the package or is available at [User Guide](@LINK_NEEDED)

@@

The function `plot_dynamics()` plots a single simulation run, as returned from the `run_simulation()` function. This function is only provided as a convenience function to provide a way to easily see the results of a simulation run. An example plot resulting from this function is shown in \@ref(fig:plot-dynamics).

@@

This is implemented in the function `run_temporal_ssfind()`, which takes a parameter object as created by the function `new_temporal_ssfind_parameter()`.

@@

## Use Cases
We will now show two use cases which illustrate applications of the package. All use cases can be used as starting points for other investigations of the behavior of the system.

The first two use cases are described in detail in the User Guide and the Partial Reproduction Vignettes. The third is taken from @REF_NEEDED for which this R package was designed. All of these use cases can be expanded to larger numbers of strains per functional group and variable values van be changed.

### Regime shifts during temporal environmental change

@@

We saved all parameter and result objects, so that all results are completely reproducible. The general flow of the experiment is identical to the one shown in \@ref(fig:runsim_example).

### The extent of hysteresis depends on community composition

One of the reasons to develop this package was to reproduce the results presented in @Bush2017. This was achieved as demonstrated in the [Partial Reproduction supplement](LINK_NEEDED). All aspects in the paper could be reproduced and are shown in the vignette.

@@

Ass all objects were saved, results can be easily reproduced and parameter can be changed to assess the impact of these changes (e.g. to conduc a sensitivity analysis).

### Effects of functional diversity on regime shifts

@@

The free and open source implementation and extension of the model used in @Bush2017 provides the means of reproducing the results published while at the same time provides the means of doing investigations which extend the original publication, as done in @REFNEEDED.

The combination of all parameter in a single parameter object as well as the return of the simulation as a result object which inherits from the parameter object and adds a results slot provides a relatively easy to use framework to implement reproducible experiments.
